---
title: "Teaching with {tidymodels}"
author: "Lisa Lendway, PhD"
institute: "Macalester College"
output:
  xaringan::moon_reader:
    df_print: paged
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      ratio: "16:9"
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE)
```

```{r library-me, echo=FALSE}
library(xaringan)
#library(xaringanExtra)
```

# Welcome!

* Find slides here: https://llendway.github.io/jsm_bof_tidymodels/
* Let's do some brief introductions: Name, pronouns, where you teach (or work, if you don't teach), and why you wanted to be here today.

---

# What is {tidymodels}?

From the Tidymodels [website](https://www.tidymodels.org/), "The tidymodels framework is a collection of packages for modeling and machine learning using tidyverse principles."

Questions:

* Have you used {tidymodels}?  
* If so, in what context did you use it?

---


# Resources

* Tidymodels [website](https://www.tidymodels.org/). 
* Tidy Modeling with R [textbook](https://www.tmwr.org/) by Julia Silge and Max Kuhn.  
* Julia Silge's [blog](https://juliasilge.com/blog/) with many examples of using tidymodels from basic to complex.
* Mine Çetinkaya-Rundel and Debbie Yuster's [slides](https://mine-cetinkaya-rundel.github.io/tidymodels-uscots-2021/tidy-up-models.html#1) from USCOTS 2021 - especially great for intro material.  
* [Example](https://advanced-ds-in-r.netlify.app/posts/2021-03-16-ml-review/) from my Advanced Data Science in R course.

---

# What we need

Here are the libraries we need (some tidyverse libraries are loaded with tidymodels so you may not need both).

```{r library-all}
library(tidyverse)     # for data visualization and wrangling
library(lubridate)     # for date wrangling
library(tidymodels)    # for modeling ... the tidy way
theme_set(theme_minimal()) # my favorite default theme
```

---

# Simple linear regression

Let's start with a simple example using the `mpg` dataset from the tidyverse library. I want to compare what I do now (on the left) with what I would need to do using tidymodels. (We can jump back to R Markdown to get a better view.)

.pull-left[
```{r}
slr <- lm(hwy ~ displ, data=mpg) 

slr %>% 
  tidy()
```

]

.pull-right[
```{r}

slr_tm <-
  # Define a linear regression model
  linear_reg() %>% 
  # Set the engine to "lm" (lm() function is used to fit model)
  set_engine("lm") %>% 
  # fit model
  fit(hwy ~ displ, data=mpg)

slr_tm %>% 
  tidy()
```

]

---


# Linear regression - prediction 

And we can do prediction. We can add `data = mpg` to `augment()` to return all original variables. And we can add, eg, `newdata = tibble(displ = c(3, 3.5))` for a couple observations.

.pull-left[

```{r}
slr %>% 
  augment()
```

]

.pull-right[

```{r}
slr_tm$fit %>% #need to pull the fit out of the rest of the model output
  augment()

slr_tm %>% 
  predict(mpg)
```

]

---

# Linear regression with more complex model 

.pull-left[

```{r}
slr_cat <- lm(hwy ~ drv*displ, data=mpg) 

slr_cat %>% 
  tidy()
```

]

.pull-right[

```{r}

cat_tm <-
  linear_reg() %>% 
  set_engine("lm") %>% 
  fit(hwy ~ drv*displ, data=mpg)

cat_tm %>% 
  tidy()
```

]
---

# Questions

* What do you like about using tidymodels right away?  
* What makes you hesitate about using tidymodels right away?  
* Other thoughts/questions?


---

# Complex model in tidymodels another way

Could be useful for easing into machine learning.

.pull-left[
```{r, eval=FALSE}
rec <- recipe(hwy ~ drv + displ, data = mpg) %>% 
  step_dummy(drv) %>% 
  step_interact(terms = ~ starts_with("drv"):displ)
  
mod <- linear_reg() %>% 
  set_engine("lm") 

workflow() %>% 
  add_recipe(rec) %>% 
  add_model(mod) %>% 
  fit(mpg) %>% # need to specify data
  tidy()
```
]
.pull-right[
```{r, echo=FALSE}
rec <- recipe(hwy ~ drv + displ, data = mpg) %>% 
  step_dummy(drv) %>% 
  step_interact(terms = ~ starts_with("drv"):displ)
  
mod <- linear_reg() %>% 
  set_engine("lm") 

workflow() %>% 
  add_recipe(rec) %>% 
  add_model(mod) %>% 
  fit(mpg) %>% # need to specify data
  tidy()
```
]

---

# Inference - using bootstrapping

When I teach inference, I start with bootstrapping to simulate the sampling distribution of the slope term. 

.pull-left[

```{r, eval=FALSE}
set.seed(155)
samples_200 <- rep_sample_n(mpg, #data we're sampling from
                            size = nrow(mpg), #size of each sample - same size
                            reps = 200, #how many samples to take
                            replace = TRUE #with or without
                                            #replacement?
                            )

samples_200 %>% 
  group_by(replicate) %>% 
  summarize(tidy(lm(hwy ~ displ))) %>% 
  filter(term == "displ") %>% 
  ggplot(aes(x = estimate)) +
  geom_histogram()
```
]
.pull-right[
```{r, echo=FALSE, fig.width=5, fig.height=5, fig.align='center'}
set.seed(155)
samples_200 <- rep_sample_n(mpg, #data we're sampling from
                            size = nrow(mpg), #size of each sample - same size
                            reps = 200, #how many samples to take
                            replace = TRUE #with or without
                                            #replacement?
                            )

samples_200 %>% 
  group_by(replicate) %>% 
  summarize(tidy(lm(hwy ~ displ))) %>% 
  filter(term == "displ") %>% 
  ggplot(aes(x = estimate)) +
  geom_histogram()
```
]

---

# Inference - bootstrapping with tidymodels

I just learned how to do this with tidymodels using this [blogpost](https://juliasilge.com/blog/superbowl-conf-int/) by Julia Silge. 

.pull-left[

```{r, eval=FALSE}
set.seed(155)

tm_intervals <- 
  reg_intervals(hwy ~ displ, # model formula
                data = mpg, # data
                model_fn = "lm", # function 
                type = "percentile", 
                times = 200, # number of samples
                keep_reps = TRUE # keep bootstrap estimates?
  )

tm_intervals %>%
  unnest(.replicates) %>%
  ggplot(aes(estimate)) +
  geom_histogram()
```
]
.pull-right[
```{r, echo=FALSE, fig.width=5, fig.height=5, fig.align='center'}
set.seed(155)

tm_intervals <- 
  reg_intervals(hwy ~ displ, # model formula
                data = mpg, # data
                model_fn = "lm", # function 
                type = "percentile", 
                times = 200, # number of samples
                keep_reps = TRUE # keep bootstrap estimates?
  )

tm_intervals %>%
  unnest(.replicates) %>%
  ggplot(aes(estimate)) +
  geom_histogram()
```
]

---

# Questions

* How does doing inference with tidymodels compare to what you usually do? 

* What other functionality would you desire? I certainly haven't covered everything.

---

# Machine learning

This is where tidymodels really excels (in my opinion). We can use similar workflows for different types of models. I am going to outline an example that is illustrated in further detail [here](https://advanced-ds-in-r.netlify.app/posts/2021-03-16-ml-review/).

First, we need a few more libraries:

```{r}
library(glmnet)            # for regularized regression, including LASSO
library(moderndive)        # for King County housing data
```

---

# tidymodels example

King Count house prices dataset from moderndive library.

```{r}
data("house_prices")

house_prices %>% 
  slice(1:5)
```


---

# tidymodels example - data splitting

Log transform the response variable price. Split the training and testing data.

```{r}
set.seed(327) #for reproducibility

house_prices <- house_prices %>% 
  mutate(log_price = log(price, base = 10)) %>% 
  select(-price)

# Randomly assigns 75% of the data to training.
house_split <- initial_split(house_prices, 
                             prop = .75)
house_training <- training(house_split)
house_testing <- testing(house_split)
```


---


# tidymodels example - recipe

Do all the pre-processing

```{r}
house_recipe <- recipe(log_price ~ ., #short-cut, . = all other vars
                       data = house_training) %>% 
  step_rm(sqft_living15, sqft_lot15) %>%
  step_log(starts_with("sqft"),
           -sqft_basement, 
           base = 10) %>% 
  step_mutate(grade = as.character(grade),
              grade = fct_relevel(
                        case_when(
                          grade %in% "1":"6"   ~ "below_average",
                          grade %in% "10":"13" ~ "high",
                          TRUE ~ grade
                        ),
                        "below_average","7","8","9","high"),
              basement = as.numeric(sqft_basement == 0),
              renovated = as.numeric(yr_renovated == 0),
              view = as.numeric(view == 0),
              waterfront = as.numeric(waterfront),
              age_at_sale = year(date) - yr_built)%>% 
  step_rm(sqft_basement, 
          yr_renovated, 
          yr_built) %>% 
  step_date(date, 
            features = "month") %>% 
  update_role(all_of(c("id",
                       "date",
                       "zipcode", 
                       "lat", 
                       "long")),
              new_role = "evaluative") %>% 
  step_dummy(all_nominal(), 
             -all_outcomes(), 
             -has_role(match = "evaluative")) %>% 
  step_normalize(all_predictors(), 
                 -all_nominal())
```


---

# tidymodels example - cv split of training

We'll use this later when we tune a parameter in the model.

```{r}
set.seed(1211) # for reproducibility
house_cv <- vfold_cv(house_training, v = 5)
```


---

# tidymodels example - set up model and workflow


```{r}
house_lasso_mod <- 
  # Define a lasso model 
  # I believe default is mixture = 1 so probably don't need 
  linear_reg(mixture = 1) %>% 
  # Set the engine to "glmnet" 
  set_engine("glmnet") %>% 
  # The parameters we will tune.
  set_args(penalty = tune()) %>% 
  # Use "regression"
  set_mode("regression")

house_lasso_wf <- 
  # Set up the workflow
  workflow() %>% 
  # Add the recipe
  add_recipe(house_recipe) %>% 
  # Add the modeling
  add_model(house_lasso_mod)
```

---

# tidymodels example - evaluate tuning parameters

Make a tuning grid (use default), fit the model using cross-validation for all `penalty_grid` values and evaluate on all the folds.

```{r}
penalty_grid <- grid_regular(penalty(),
                             levels = 20)
house_lasso_tune <- 
  house_lasso_wf %>% 
  tune_grid(
    resamples = house_cv,
    grid = penalty_grid
    )

# rmse averaged over all folds:
house_lasso_tune %>% 
  collect_metrics() %>% 
  filter(.metric == "rmse")
```

---

# tidymodels example - fit final model

We choose the best penalty parameter as the one with the smallest cross-validated RMSE. We then use that parameter to fit the model to the training data.

```{r}
best_param <- house_lasso_tune %>% 
  select_best(metric = "rmse")

house_lasso_final_wf <- house_lasso_wf %>% 
  finalize_workflow(best_param)

house_lasso_final_mod <- house_lasso_final_wf %>% 
  fit(data = house_training)

house_lasso_final_mod %>% 
  pull_workflow_fit() %>% 
  tidy() 
```


---


# Open discussion

* What do you like about tidymodels?  
* What reservations do you have about tidymodels?  
* Where in the statistics/data science curriculum do you think is the best place to introduce tidymodels? And why?  
* What didn't we talk about that we should?  



