---
title: "Teaching with {tidymodels}"
author: "Lisa Lendway, PhD"
institute: "Macalester College"
output:
  xaringan::moon_reader:
    df_print: paged
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE)
```

```{r library-me, echo=FALSE}
library(xaringan)
#library(xaringanExtra)
```

# What is {tidymodels}?

From the Tidymodels [website](https://www.tidymodels.org/), "The tidymodels framework is a collection of packages for modeling and machine learning using tidyverse principles."

Questions:

* Have you used {tidymodels}?  
* If so, in what context did you use it?

---


# Resources

* Tidymodels [website](https://www.tidymodels.org/). 
* Tidy Modeling with R [textbook](https://www.tmwr.org/) by Julia Silge and Max Kuhn.  
* Julia Silge's [blog](https://juliasilge.com/blog/) with many examples of using tidymodels from basic to complex.
* Mine Çetinkaya-Rundel and Debbie Yuster's [slides](https://mine-cetinkaya-rundel.github.io/tidymodels-uscots-2021/tidy-up-models.html#1) from USCOTS 2021 - especially great for intro material.

---

# What we need

Here are the libraries we need (some tidyverse libraries are loaded with tidymodels so you may not need both).

```{r library-all}
library(tidyverse)     # for data visualization and wrangling
library(lubridate)     # for date wrangling
library(tidymodels)    # for modeling ... the tidy way
```

---

# Simple linear regression

Let's start with a simple example using the `mpg` dataset from the tidyverse library. I want to compare what I do now (on the left) with what I would need to do using tidymodels. (We can jump back to R Markdown to get a better view.)

.pull-left[

## `lm()`

```{r}
slr <- lm(hwy ~ displ, data=mpg) 

slr %>% 
  tidy()
```

]

.pull-right[

## tidymodels

```{r}

slr_tm <-
  # Define a linear regression model
  linear_reg() %>% 
  # Set the engine to "lm" (lm() function is used to fit model)
  set_engine("lm") %>% 
  # fit model
  fit(hwy ~ displ, data=mpg)

slr_tm %>% 
  tidy()
```

]

---

# Linear regression - prediction 



---

# Linear regression with categorical predictor 

Adding a categorical variable.

---

# Linear regression with interaction

---

# Simulation

---

# Machine learning

---


